Proof of uniform convergence for a cell-centered AP discretization of the hyperbolic heat equation on general meshes

### 2015-05-19 Answers_for_reviewers.tex

\title{Answers for the Reviewers  for the paper "Proof of uniform convergence for a cell-centered AP discretization of the hyperbolic heat equation on general meshes"}

\author{Emmanuel Franck\footnotemark[1] \footnotemark[2],
C. Buet \footnotemark[3], 
B. Despr\'es\footnotemark[4], 
T. Leroy\footnotemark[4] \footnotemark[3]
}
 \footnotetext[1]{INRIA Nancy Grand-Est and IRMA Strasbourg}
 \footnotetext[3]{CEA, DAM, DIF, F-91297 Arpajon Cedex,France}
 \footnotetext[4]{Laboratoire Jacques-Louis Lions,
  Universit\'e Pierre et Marie Curie,
  75252 Paris Cedex 05,
  France}
\footnotetext[2]{Email adress: emmanuel.franck@inria.fr}

\maketitle

\section{Work done}

The authors deeply thank the referees for their reviews.
We understant that our work is judged interesting and worth of publications,
but that the a more detailed of explanation of the estimates and a proof of the fully 
discrete scheme were necessary.
We have done this work, together with a reexamination line after line of all estimates.

Our answer to the all question raised in the reports is below.
In {\bf bold} are the questions/remarks, our answers follow immediately.

\section{Answers for the report 2M}

\begin{itemize}
\item {\bf A massive overuse of "Constant C" and the issue of the estimates' time-dependence \dots}

In order to give a positive answer to the question of the "massive overuse of the constant $C$", we have
completely re-examine all that concerns the constants.
In order to clarify, we have adopted the following strategy:

a)  the four main constants which correspond to the four branches of the AP diagram and the new AP diagram
are denoted as $~_\downarrow C$, $C^\rightarrow$, $C_\downarrow$ and $C_\leftarrow$.
Moreover the dependence with respect to the constant which measures the quality
of the mesh, denoted as $C_\mathcal M$, and with respect to the Gronwall exponentials
is indicated systematically {\bf in dimension $n=1$}.
In 1D, it is now clear that the Gronwall exponentials have zero dependence with respect to $1/\varepsilon$.
So they are perfectly bounded for a given time $T$. There is no exponential of exponential.

In dimension two, we do not indicate the extra dependance 
with respect to $C_\mathcal M$ and  the Gronwall-exponentials,
since it would make the reading much heavier
once again, and  it would bring almost zero with respect to the 1D case.

The constant $S$ in section 3.5 is now renamed $E_3$.

b) the "sub-sectioning" tries to be more rigorous. For example 
a new subsection 3.4 now explains the various norms and approximation/interpolation estimates,
subsection 3.5 is split into sub subsections entitled stability, consistency, ..., 
subsection 3.6 is split into a rescaling sub subsection, and so on.

c) The main inequality (one of the four branch) of the AP diagram yields a lemma
which is written at the beginning of the main subsections (3.5, 3.6, 3.7).

d) A new section (asked by the other referred) explains how to use a simple method
(from Despres, Math of Comp 04) to obtained a $\sqrt{\Delta t}$ error term in time, uniformly
with respect to $h$ and $\varepsilon$.
It allows to formulate the  theorem 1.1 immediately in the introduction.

In any cases, we mention to the referee that a huge  work has been done to fulfill his expections.
It also seems to us more rigorous now.

\begin{itemize}

\item {\bf Page 7, Proposition 2.2: as far as I understand, the proof yields C = T. So there's no need for this first constant. Agree ? Same for Prop. 3.1 ?}

we have corrected the constant in the proposition 2.2 page 7

\item {\bf The former comment may seem anecdotal, but it reveals (at least) 2 interesting things: }

\begin{itemize}
\item 
{\bf 1.
assuming p smooth, the approximation (on continuous solutions) is reliable mainly for $t <<\frac1 \epsilon$ }

\item {\bf 2. all the remaining Propositions (pages 11-14) ask for the Gronwall lemma in their proofs: hence we obtain bounds growing exponentially in time, suggesting that numerical approximations (discrete solutions) are now r \dots}

\end{itemize}

{\bf Hence I think that the authors shouldn't try to hide this apparent discrepancy under the carpet, because it may be checked numerically rather easily: it suffices to run the code which produced the Figure 6 several times, and scrutinize whether we see a growth in time which is linear, polynomial or exponential. Moreover, this will give valuable information on the sharpness of the error bounds rigorously proven}

Already answered: the  important parameters of the  constants are now fully detailed in dimension 1, and partially
in dimension 2.

\item {\bf Page 8, the estimate (27): the proof is quite easy, so why not displaying it in order to wipe off the C ? It suffices to set up a classical computation, from e.g. DiPerna,
$$
\frac12 \frac d {dt} \int_R p^2 dx + \frac1\sigma \int_R |\partial_x p |^2\leq 0 \Longrightarrow \|  p(t,\cdot) \|_{L^2}\leq  \|  p_0 \|_{L^2}.
$$
which yields, without introducing any C,
$$
\int_0^T \int_R |\partial_x p |^2 dt \leq \frac \sigma 2 \|  p_0 \|_{L^2}^2.
$$
}

Corrected.

\item {\bf Page 11, proof of (32)-(33)}

\begin{itemize}

\item {\bf 1.  one can find a constant $C \geq 0$ such that the next term $[\dots]$. Is this
C the one of the grid introduced in Hypothesis 2.1 ?}

It is the mesh constant, now explicitly given as $C_\mathcal M$ everywhere.

\item {\bf 2. The last term in (28) $[\dots]$: seems like it's (29) instead }

(28)$\to$(29), corrected.

\item {\bf 3.  using a Cauchy-Schwarz inequality $[\dots]$: in my opinion, one should invoke instead Jensen's inequality. Indeed \dots}

We have change ``the Cauchy-Schwarz inequality'' to ``the Jensen inequality''.

\end{itemize}

\item {\bf Pages 11-12, in Proposition 2.6, the C(T) is an exponential of the time T: it must be indicated and the discrepancy with respect to the linear growth found in Proposition 2.2 should be commented. Moreover, concerning the estimate (35): the statement is correct, except that it would require a little bit of explanations, as it relies on a bootstrap-type argument based on estimate (27), which wasn?t stated really correctly (as already seen).}

Pages 11-12: all the constants have been given.

\item 
{\bf Pages 20-27: the statements and proofs of Props. 3.4 and 3.5 are A MESS!}

Entirely reshaped as already mentioned above.

\begin{itemize}
\item

{\bf 1. When studying the different terms in $E'(t)$, please use either itemize or enumerate, and perhaps forward heavy calculations to an 
Appendix in order to lighten stuff,}

Done.

\item {\bf 2. Props. 3.6-8 are indeed "technical Lemmas" and should be stated accordingly (cf. comments on the 1D case of Prop. 2.8). Their proofs might be forwarded too.}

\item {\bf 3. the 1D comment about Jensen's inequality appears to apply here again, bottom of page 26 }

Jensen does not show up anymore.

\end{itemize}

\item {\bf Page 36, 4: perhaps a bibliographic reference for the solution of the telegraph's equation may be useful to the reader. Moreover, there is a conflict of notation between the $\gamma$ used here and the one invoked in the proofs on pages 25-26-27}

Page 36 : The conflict for the notation have been corrected. For the solution there are not reference, we have construct this solution using a separation of the variable. we have added the details.

\item {\bf As said before, there's a question which should be addressed, that is the rate of amplification in time of the errors: the Gronwall?s Lemma produces a time-exponential rate, so it's interesting to check out numerically if this feature is a discrepancy of the analysis, or if it holds practically. Such a point was addressed, in a totally different context of scalar balance laws.

{\it D. Amadori, L. Gosse, Transient $L1$
 error estimates for well-balanced schemes on non-resonant scalar balance laws, J. Differential Equations 255 (2013) 469-502}

}

\item Last point: in 1D we have explicit all the constant. We obtain time-exponential rate as you expect. Numerically we constat a a time-linear rate, we have added a remark in the numerical part at this subject.

\item 
Additional modifications: Page 5 we have modified the constant for the 1D regularity mesh;
We have correct the constant in the proposition 2.5 page 8.
\end{itemize}

\end{itemize}

\begin{itemize}

\item Comments and suggestions
\begin{itemize}
\item 
{\bf Pages 2-': with an Introduction that long, it makes sense to insert "sub- sections" entitled "Precisions on AP discretizations" (right after Remark 1.2), and ?Organization of the paper? (right after the proof of Prop 1.3 }

 We have added two subsections : "Precisions on AP discretizations" and "Organisation of the proof"

\item {\bf Page 5: is there a connection between "nodal-based" FV schemes (advocated in the Intro.) and the classical Godunov scheme where 2D Riemann problems are to be solved at each node of the computational grid ? See for instance the elementary case of a non-resonant scalar balance law in,
L. Gosse, A two-dimensional version of the Godunov scheme for scalar balance laws. SIAM J. Numer. Anal. 52 (2014) 626-652.
If so, the sentence in the Introduction "it strongly questions ..." wouldn't be that surprising. Moreover, on page 27, the authors speak about the "incorporation in the approximate nodal Riemann solver", thus reinforcing this vague impression of similarity with 2D Riemann solvers}

We  added 2 sentences about the 2D Riemann problem and the nodal scheme. 
We agree that there is a similarity, but essentially concerning the final formula itself.
The method is rather different.

\item {\bf Page 6: The big footnote is confusing: in my opinion, it would be better to state clearly that, in order to build a discretization able to produce a 2D AP numerical process, the authors choose to renounce to the well- balanced property of the original 1D Gosse-Toscani scheme (except on uniform Cartesian grids) while retaining only its AP character. This ex- plains why the algebraic expressions differ from each other.}

We have added a sentence to explain why we use this formulation of the scheme and not the classical Gosse-Toscani scheme.

For the remark about " the authors choose to renounce to the well- balanced property of the original 1D Gosse-Toscani scheme (except on uniform Cartesian grids) while retaining only its AP character" it is not clear that we say that. Indeed the scheme is well balanced for a subset of steady states (see annex of the paper "design of AP scheme ..."): the steady state associated to $u=Cts$ and $p$ linear. This property is perhaps impotent for the uniform convergence. At the end we renounce only to the steady states associated to $\nabla{u}=\nabla\times \phi$ with $\phi$ a potential.

\item {\bf Page 7, Proof of Prop. 2.2, the notation choice $u^\epsilon=-\frac\sigma \epsilon \partial _x p^\epsilon$  is confusing:
according to (13), a notation like $v^\epsilon$ looks better.}

 We have correct this part introducing $v$ to obtain a not confusing proof.

\item {\bf Page 10, last lines, what is the definition of $\mathbf u^{\eps}$ ? This notation doesn?t
appear previously.}

 We have replaced is $\mathbf{u}^{\eps}$ by $u^{\eps}$

\item {\bf Page 14, the Proposition 2.8 is indeed a (technical) Lemma, and this should be stated accordingly}

We have replaced the proposition 2.8 by the Lemma 2.8

\item {\bf Page 14, end of the 1D section: according to what is displayed numerically in the §4, Figure 6, we see a second-order accuracy
 (in h) for $\epsilon \approx  h^2$. This appears to be an extension in 2D of an observation made in a paper devoted to 
 "TAHO schemes", see especially §7.1.2 in,
D. Aregba, M. Briani, R. Natalini, Time Asymptotic High Order Schemes for Dissipative BGK Hyperbolic Systems, arXiv:1207.6279}

We have added the reference on the paper "Time asymptotic high order ..." at the end of the 2D in the numerical results.

\item {\bf Page 15, Proposition 3.1: same comment than for Proposition 2.2}

We have corrected the proposition

\item {\bf Page 15, eqn. (45), there is a "(" missing in the first line.}

We have added the missing "(".
\end{itemize}

\item Minor issues and Typos
\begin{itemize}
\item 1: We have replaced $P^0$ by "first order"
\item 2 : we have corrected this sentence 
\item 3 : we have corrected this sentence 
\item 4 : we have corrected this sentence. (17) correspond to the scheme not the smooth data.
\item 5 : we have corrected this sentence. 
\item 6 : we have corrected your two remarks in this page 
\item 7 : we have corrected this sentence. 
\item 8: We have replaced $\gamma$ by $\tau$ in the numeric part.
\item 9: The date for the L. Gosse paper have been modified. for the S. Jin article we have copy the reference written on the website of S. Jin 
\end{itemize}
\end{itemize}

\section{Answers for the report 2C}

Before answering to explain what we have done, we would like to make a general comment.

We still consider that the space discretization is the most important, in the sense
that such multiD method can be used by researchers and engineers only if they are proved
to be convergent, in the usual sense.
That it is converge for a family of smooth enough functions.
The question of the boundary in time
layer plays little role for the application we have in mind.
On the contrary, discontinuity of the coefficients of the equations may generate 
some boundary layers in space. For radiation equation this is called the Milne problem.
This issue is of tremendous importance, but seems to us difficult to address in general, not only for the scheme that we develop.

On the other hand our method is the first which is proved to be convergent: what we think is of real importance
is the fact that usual Finite Volume technique through the edges are not able have a geometrical
obstruction for AP on a general mesh. This is why our method for which fluxes are corner based 
seems a real advance.

Now we answer to 
{\bf  The issue concerning the time
discretization must be addressed correctly.
}.

This issue was not addressed at all because it seems to us that the convergence in space was the real point.
Nevertheless we do agree with the referee.
This is why was have now done the following.

\begin{itemize}
\item the theorem of convergence for a fully explicit scheme is written in the introduction, and proved
in section  4b by the abstract method (the main idea was given in  despres 04, but
the new proof is much more direct). This method easily give a uniform estimate 
between the continuous and fully discrete solutions, so proves the theorem when combined with the previous results.

\item Furthermore we have restructure most of the material. The proof of the naive estimate in 2D 
is in two parts, the technical details being in the appendix is asked by the other referee.
We also give a much more detailed presentation of the constants
$~_\downarrow C$, $C^\rightarrow$, $C_\downarrow$ and $C_\leftarrow$ with the regularity of the initial data
and other constants.

\item We feel the entire work is more complete rigorous with these two modifications.

\end{itemize}

The proposed method has already been applied to to a non linear diffusion equation in  (see references):
{\it C. Buet, B. Despr\'es, E. Franck, \emph{An asymptotic
preserving scheme with the maximum principle for the $M_1$ model on distorted
meshes}, C.R. Acad. Sci., Paris, S\'er. I, Math., Vol 350,  11-12
pp 633-638,  2012.}

The propose references are relevant. They have been added and discussed with the other references in the introduction.

\end{document}

### 2015-05-19 Corrected_version_2.tex
modif article

\documentclass[a4paper,french,english,10pt]{article}
\usepackage{a4wide}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{lmodern}
%\usepackage{ae}

@@

\newcommand\pej{\bar{p}_j}
\newcommand\uej{\bar{\uu}_j}
\newcommand\uer{\bar{\uu}_r}

@@

\newcommand\cj{\cc_j}
\newcommand\br{\bb_r}
\newcommand\diam{\mbox{diam}}

@@

\section{Introduction}

We address the convergence analysis  on unstructured meshes
of diffusion
asymptotic preserving schemes for the discretization
of a problem with a stiff parameter denoted as $\eps\leq 1$.
The model problem considered in this work
is  the hyperbolic heat equation in the domain 
$t\geq 0$ and  $x\in \Omega\subset \mathbb R^n$
\begin{equation} \label{eq:mod1}
P^\varepsilon: \qquad 
\left\{\begin{array}{ll}
\ds\partial_t p^{\eps}+\frac{1}{\eps}\operatorname{div}(\uu^{\eps})=0,
& p^{\eps}\in \mathbb R, \\
\\
\ds\partial_t \uu^{\eps}+\frac{1}{\eps}\nabla
p^{\eps}=-\frac{\sigma}{\eps^2}\uu^{\eps},
& \uu^{\eps}  \in \mathbb R^n
 \end{array}\right.
\end{equation}
discretized with first order finite volume schemes.
This problem is representative
of many  transport problem such
as transfer and neutron transport,  for which the small parameter
$\eps$ is the ratio of two very different sound velocities and
$\sigma$ is the absorption or the opacity.
For simplicity both $\eps$ and  $\sigma>0$ are kept constant in space in this study.
The system (\ref{eq:mod1}) can also be introduced as a specific linearization
of a pressure-velocity system of partial differential equations
in the acoustic regime. In this work we will need the following well known energy estimates concerning 
the Cauchy problem for the partial differential equation (\ref{eq:mod1}): 
\begin{proposition}\label{bee}
If $\Omega= \mathbb R^n$ or $\Omega= \mathbb T^n$ and if the initial data $\V_0=(p_0,\uu_0) \in H^p( \Omega)$ then
\begin{equation}\label{bee1}
\vert\vert \V \vert \vert_{H^p( \Omega)}\leq \vert\vert \V_0 \vert \vert_{H^p( \Omega)}
\end{equation}
 and moreover 
\begin{equation}\label{bee2}
\frac{\sigma}{\eps^2}\vert\vert \uu \vert \vert^2_{L^2([0,T];H^p( \Omega))}\leq \vert\vert \V_0 \vert \vert^2_{H^p( \Omega)}.
\end{equation}
\end{proposition}
We will consider well prepared data in the sense that:
 $p^\eps(t=0)$
is independent of $\eps$ and is sufficiently smooth;
the initial velocity satisfies the equality in the second equation
of (\ref{eq:mod1}) at leading order. It writes 
\begin{equation} \label{eq:iniwp}
p^\eps(t=0)=p_0 \mbox{ and } 
\mathbf u^\eps(0)=-\frac\eps\sigma \nabla p_0.
\end{equation}
For such well prepared data, it can be easily shown that
the formal limit of $P^\varepsilon$ for small $\eps$ is
\begin{equation} \label{eq:mod2}
P^0: \qquad
\partial_t p - \frac1\sigma \Delta p=0. 
\end{equation}

\begin{remark}\label{sigmanotzero}
We do not consider the regime $\sigma \rightarrow 0$, since it introduces a singularity both
  in the initial data of the hyperbolic heat equation and in the limit parabolic equation.
\end{remark}
\subsection{Precision of AP discretizations}
 Before addressing the main difficulty of this work  which is the discretization on  
  unstructured meshes, we briefly recall 
the  now well known notion 
of an asymptotic preserving technique \cite{jinbase}-\cite{jinreview}
which  is illustrated in the figure \ref{fig:diag1}.
For the simplicity of the presentation,
we will consider mainly semi-discrete numerical methods, this is why
the time step does not show up.
In  figure \ref{fig:diag1} the parameter $h$ designs a numerical method
with characteristic length $h\leq 1$:
so we assume  a numerical method $P_h^\varepsilon$ for the discretization
of $P^\varepsilon$.  
\begin{definition}[Uniform AP] \label{defAPu}
If  $P_h^\varepsilon$ is consistent with $P^\varepsilon$
uniformly with respect to $\varepsilon$, 
then we  say that the  scheme $P_h^\varepsilon$
is  uniformly AP  (uniformly asymptotic preserving).  
\end{definition}

@@

This property is simpler to analyze than the uniform AP. It explains
why it has been very fruitful in the past.
In 1D, many AP schemes have been designed for some PDE and physical problems: S.
Jin, C. D. Levermore \cite{jinlev} or L. Gosse, G. Toscani \cite{Gosse} for the
hyperbolic heat equation, M. Lemou, L. Mieussens, N. Crouseilles
\cite{lemou}-\cite{MMvlasov}-\cite{couplingRad} for some kinetic equations, L.
Gosse \cite{GosseSn}, C. Buet and co-workers \cite{buet} or S. Jin and C. D.
Levermore \cite{SnJin} for $S_N$ equations and C. Berthon, R. Turpault
\cite{ber1}-\cite{ber3}-\cite{ber4} for generic systems and a non linear
radiative transfer model. Recently some asymptotic preserving schemes for linear
systems and non linear radiative transfer model have been designed in 2D
\cite{glaceap}-\cite{cras}-\cite{FVCA6}. However for this type of schemes it is
 difficult to obtain convergence estimates due to the competition between
the two parameters $\eps$ and $h$. To our knowledge this type of
proof are only given for uniform grids \cite{glaceap} (consistence and
stability, Lax theorem), \cite{Gosse} ($L^1$ and BV estimates), \cite{mmcv}
($L^2$ estimates).
The goal of this work is to prove the uniform AP property on unstructured 
grids.

To this end we adapt a strategy developed in 
 \cite{cveps} in a slightly different context.
It relies on the derivation of a priori estimates attached to the AP diagram in figure
\ref{fig:diag1}. To have a more global perspective
on this strategy,  let us assume 
some natural abstract a priori  estimates for a given norm
which is in our work based on   $\|f\|=\|f\|_{L^2([0,T]\times \Omega ) }$
where $T>0$ is a given final time, $\Omega=\mathbb{R}$, in 1D or $\Omega=[0,1]^2$ with periodic boundary conditions in 2D.
We assume four positive constants $a,b,c,d>0$ and another universal constant
 $C$ such that 
\begin{equation}\label{eq:mod3}
\| P^\varepsilon - P^0 \| \leq
C \varepsilon ^{a},
\end{equation}
\begin{equation}\label{eq:mod7}
\| P_h^\varepsilon - P^\varepsilon   \|_{ {\rm naive}} \leq
C \varepsilon ^{-b}h^c,
\end{equation}
\begin{equation}\label{eq:mod4}
\| P_h^0 - P^0   \| \leq
C h^d.
\end{equation}
The first inequality expresses that $P^0$ is indeed the limit of
$P^\varepsilon$. 
The second inequality is typically
based on  non AP error bounds. This is why we refer
to it as the naive error bound. 
The third inequality is the AP property.
A fourth inequality for
$\| P_h^\eps-P_h^0\|$ 
is of course
required to close the diagram.
We assume that it can be obtained in a form similar to 
(\ref{eq:mod3})
\begin{equation}\label{eq:mod3-bis}
\| P^\varepsilon_h - P^0_h \| \leq
C  \varepsilon ^{e}, \qquad e>0.
\end{equation}

\begin{pro} \label{prop:1}
 Assume that all these inequalities are at 
hand and that $d\geq c$ and $e\geq a$.
 Then the uniform AP holds with a rate at least
$O\left(h{^\frac{ac}{a+b}  } \right)$.
\end{pro}
\begin{proof}
The triangular inequality writes
$$
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
\min
\left(
\| P_h^\varepsilon - P^\varepsilon   \|_{ {\rm naive}},
\| P_h^\eps-P_h^0\|+
\| P_h^0 - P^0   \| +
\| P^\varepsilon - P^0 \|
\right)
$$
which yields, using $\min(x,y+z)\leq \min(x,y)+\min(x,z)$ and   $e\geq a$.
\begin{equation} \label{eq:toutdebut}
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
C\left(
\min
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right)+h^d+
\min\left(
 \varepsilon ^{-b}h^c,
\eps^e
\right)\right)
\leq C\left(
2\min  
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right) +h^d\right).
\end{equation}

We define a  threshold value $\varepsilon_{\rm thresh} $
by  $\varepsilon_{\rm thresh}  ^{-b}h^c=
\eps_{\rm thresh}^a$. So either $\eps\leq \varepsilon_{\rm thresh} $
so that $\min
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right)\leq \varepsilon_{\rm thresh}^a =h^{\frac{ac}{a+b}  }$, or 
$\eps\geq \varepsilon_{\rm thresh} $ and the same bound is obtained by 
taking the other term as the minimum.
And since $d\geq c$, one gets the abstract bound
$
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
3Ch^{\frac{ac}{a+b}  }
$ which ends the proof.
\end{proof}

@@

@@

\begin{hyp}[Regularity of the mesh in 1D] \label{geometrie1d}
 We consider that there exists 
a universal constant $C_{\mathcal{M}}>0$ independent of the mesh size
$h=\sup_{j\in \mathbb Z} \Delta x_j $
so that one has 
$$
C_{\mathcal{M}} h \leq \Delta x_j \leq h \qquad \forall j \in \mathbb{Z}.
$$
\end{hyp}

@@

These estimates characterize the dissipation rate 
of both schemes.

@@

In order to establish this estimate, we will use the method introduced by C.

@@

that we consider  is

>>>>

\documentclass[a4paper,french,english,10pt]{article}
\usepackage{a4wide}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{lmodern}

\usepackage{picture}
%\usepackage{ae}

@@

\newcommand\pej{\overline{p}_j}
\newcommand\uej{\overline{\uu}_j}
\newcommand\uer{\overline{\uu}_r}

@@

\newcommand\cj{\cc_j}
\newcommand\ccr{\cc_r}
\newcommand\br{\bb_r}
\newcommand\bj{\bb_j}

\newcommand\diam{\mbox{diam}}

@@

\section{Introduction}
%
%\setlength{\unitlength}{.3cm}
%\begin{picture}(1, 1) 
%\thicklines
%\put(0,0){\vector(0,1){1}}
%\put(0,1){\line(1,0){1}}
%\put(1,1){\line(0,-1){1}}
%\put(1,0){\line(-1,0){1}}
% \end{picture}
%
%
%$$
%{\setlength{\unitlength}{.3cm}
%\begin{picture}(1, 1) 
%%\thicklines
%\put(0,0){\vector(0,1){1}}
%\put(0,1){\line(1,0){1}}
%\put(1,1){\line(0,-1){1}}
%\put(1,0){\line(-1,0){1}}
% \end{picture}
%} 
%$$
%
%$$
%C^{\begin{picture}(1, 1) 
%%\thicklines
%\put(1,1){\vector(1,0){15}}
%\end{picture}
%}
%$$
%$$
%C^\Longrightarrow
%$$
%

We address the convergence analysis  on unstructured meshes
of diffusion
asymptotic preserving schemes for the discretization
of a problem with a stiff parameter denoted as $\eps\leq 1$.
The model problem considered in this work
is  the hyperbolic heat equation in the domain 
$t\geq 0$ and  $x\in \Omega\subset \mathbb R^n$
\begin{equation} \label{eq:mod1}
P^\varepsilon: \qquad 
\left\{\begin{array}{ll}
\ds\partial_t p^{\eps}+\frac{1}{\eps}\operatorname{div}(\uu^{\eps})=0,
& p^{\eps}\in \mathbb R, \\
\\
\ds\partial_t \uu^{\eps}+\frac{1}{\eps}\nabla
p^{\eps}=-\frac{\sigma}{\eps^2}\uu^{\eps},
& \uu^{\eps}  \in \mathbb R^n
 \end{array}\right.
\end{equation}
discretized with first order finite volume schemes.
This problem is representative
of many  transport problem such
as transfer and neutron transport,  for which the small parameter
$\eps$ is the ratio of two very different sound velocities and
$\sigma$ is the absorption or the opacity.
For simplicity both $\eps$ and  $\sigma>0$ are kept constant in space in this study.
The system (\ref{eq:mod1}) can also be introduced as a specific linearization
of a pressure-velocity system of partial differential equations
in the acoustic regime. In this work we will need the following well known energy estimates concerning 
the solution $\V^\eps$ of the Cauchy problem for the partial differential equation (\ref{eq:mod1}).

\begin{proposition}\label{bee}
If $\Omega= \mathbb R^n$ or $\Omega= \mathbb T^n$, then
\begin{equation}\label{bee1}
\vert\vert \V^\varepsilon \vert \vert_{H^p( \Omega)}\leq \vert\vert \V^\varepsilon(0) \vert \vert_{H^p( \Omega)}
\end{equation}
 and moreover 
\begin{equation}\label{bee2}
\frac{\sigma}{\eps^2}\vert\vert \uu^\varepsilon \vert \vert^2_{L^2([0,T];H^p( \Omega))}\leq \vert\vert \V^\varepsilon(0) \vert \vert^2_{H^p( \Omega)}.
\end{equation}
\end{proposition}

We will consider well prepared data in the sense that:
 $p^\eps(t=0)$
is independent of $\eps$ and is sufficiently smooth;
the initial velocity satisfies the equality in the second equation
of (\ref{eq:mod1}) at leading order. It writes 
\begin{equation} \label{eq:iniwp}
p^\eps(t=0)=p_0 \mbox{ and } 
\mathbf u^\eps(0)=-\frac\eps\sigma \nabla p_0.
\end{equation}
For such well prepared data, it can be easily shown that
the formal limit of $P^\varepsilon$ for small $\eps$ is
\begin{equation} \label{eq:mod2}
P^0: \qquad
\partial_t p - \frac1\sigma \Delta p=0. 
\end{equation}

\begin{remark}\label{sigmanotzero}
We do not consider the regime $\sigma \rightarrow 0$, since it introduces a singularity both
  in the initial data of the hyperbolic heat equation and in the limit parabolic equation.
\end{remark}

\subsection{Precision of AP discretizations}

 Before addressing the main difficulty of this work  which is the discretization on  
  unstructured meshes, we briefly recall 
the  now well known notion 
of an asymptotic preserving technique \cite{jinbase}-\cite{jinreview}
which  is illustrated in the figure \ref{fig:diag1}.
For the simplicity of the presentation,
we will consider mainly semi-discrete numerical methods, this is why
the time step does not show up in the graphic.
The parameter $h$ designs a numerical method
with characteristic length $h\leq 1$:
that is  we assume  a numerical method $P_h^\varepsilon$ for the discretization
of $P^\varepsilon$.  
\begin{definition}[Uniform AP] \label{defAPu}
If  $P_h^\varepsilon$ is consistent with $P^\varepsilon$
uniformly with respect to $\varepsilon$, 
then we  say that the  scheme $P_h^\varepsilon$
is  uniformly AP  (uniformly asymptotic preserving).  
\end{definition}

@@

This property is simpler to analyze than the uniform AP. It explains
why it has been very fruitful in the past.
In 1D, many AP schemes have been designed for some PDE and physical problems: S.
Jin, C. D. Levermore \cite{jinlev} or L. Gosse, G. Toscani \cite{Gosse} for the
hyperbolic heat equation, M. Lemou, L. Mieussens, N. Crouseilles
\cite{lemou}-\cite{MMvlasov}-\cite{couplingRad} for some kinetic equations, L.
Gosse \cite{GosseSn}, C. Buet and co-workers \cite{buet} or S. Jin and C. D.
Levermore \cite{SnJin} for $S_N$ equations and C. Berthon, R. Turpault
\cite{ber1}-\cite{ber3}-\cite{ber4} for generic systems and a non linear
radiative transfer model. Recently some asymptotic preserving schemes for linear
systems and non linear radiative transfer model have been designed in 2D
\cite{glaceap}-\cite{cras}-\cite{FVCA6}.
Other application to non linear hyperbolic systems of conservation laws with stiff diffusive relaxation is to be found is
\cite{F1}. Relaxation systems are treated in \cite{F2}. More general situation 
for transport and discrete velocity systems are in \cite{F3,F4}.
 However for this type of schemes it is
 difficult to obtain convergence estimates due to the competition between
the two parameters $\eps$ and $h$. To our knowledge this type of
proof are only given for uniform grids \cite{glaceap} (consistence and
stability, Lax theorem), \cite{Gosse} ($L^1$ and BV estimates), \cite{mmcv}
($L^2$ estimates).
The goal of this work is to prove the uniform AP property on unstructured 
grids.

To this end we adapt a strategy developed in 
 \cite{cveps} in a slightly different context.
It relies on the derivation of a priori estimates attached to the AP diagram in figure
\ref{fig:diag1}. To have a more global perspective
on this strategy,  let us assume 
some natural abstract a priori  estimates for a given norm
which is in our work based on   $\|f\|=\|f\|_{L^2([0,T]\times \Omega ) }$
where $T>0$ is a given final time, $\Omega=\mathbb{R}$, in 1D or $\Omega=[0,1]^2$ with periodic boundary conditions in 2D.
We assume four  constants $a,b,c,d>0$ and four  additional  constants $~_\downarrow C, C ^\rightarrow  ,C_\leftarrow,  C_\downarrow>0>0$ 
 such that the error attached to the branches of the AP diagram can be bounded like  
 \begin{equation}\label{eq:mod7}
\| P_h^\varepsilon - P^\varepsilon   \|_{ {\rm naive}} \leq~_\downarrow C \varepsilon ^{-b}h^c, % \qquad ~_\downarrow C>0,
\end{equation}
\begin{equation}\label{eq:mod3-bis}
\| P^\varepsilon_h - P^0_h \| \leq
C ^\rightarrow  \varepsilon ^{e}. % \qquad C ^\rightarrow  >0.
\end{equation}
\begin{equation}\label{eq:mod4}
\| P_h^0 - P^0   \| \leq
C_\downarrow  h^d,  % \qquad C_\downarrow>0.
\end{equation}
\begin{equation}\label{eq:mod3}
\| P^\varepsilon - P^0 \| \leq
C_\leftarrow  \varepsilon ^{a},     %\qquad C_\leftarrow>0,
\end{equation}
 The first inequality is  the naive error bound which typically blows up for small $\varepsilon$.
The second inequality 
for
$\| P_h^\eps-P_h^0\|$ 
is 
assumed to have  a form similar to the last one which  expresses that $P^0$ is  the limit of
$P^\varepsilon$. 
The third inequality corresponds to  the usual AP property.

\begin{pro} \label{prop:1}
 Assume that all these inequalities are at 
hand and that $d\geq c$ and $e\geq a$.
 Then the uniform AP holds with a rate at least
$O\left(h{^\frac{ac}{a+b}  } \right)$.
\end{pro}

\begin{proof}
The triangular inequality writes
$$
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
\min
\left(
\| P_h^\varepsilon - P^\varepsilon   \|_{ {\rm naive}},
\| P_h^\eps-P_h^0\|+
\| P_h^0 - P^0   \| +
\| P^\varepsilon - P^0 \|
\right)
$$
which yields, using $\min(x,y+z)\leq \min(x,y)+\min(x,z)$, $d\geq c$  and   $e\geq a$, 
\begin{equation} \label{eq:toutdebut}
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
C\left(
\min\left(
 \varepsilon ^{-b}h^c,
\eps^e
\right)+h^d+\min
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right)
\right)
\leq C\left(
2\min  
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right) +h^d\right)
\end{equation}
with $C=\max\left( _\downarrow C,C^\rightarrow, C_\downarrow  ,  C_\leftarrow \right)$.
We define a  threshold value $\varepsilon_{\rm thresh} $
by  $\varepsilon_{\rm thresh}  ^{-b}h^c=
\eps_{\rm thresh}^a$. So either $\eps\leq \varepsilon_{\rm thresh} $
so that 
$$
\min
\left(
 \varepsilon ^{-b}h^c,
\eps^a
\right)\leq \varepsilon_{\rm thresh}^a =h^{\frac{ac}{a+b}  },
$$ 
or 
$\eps\geq \varepsilon_{\rm thresh} $ and the same bound is obtained by 
taking the other term as the minimum.
And since $d\geq c$, one gets the abstract bound
$
\| P_h^\varepsilon - P^\varepsilon   \|
\leq
3Ch^{\frac{ac}{a+b}  }
$ which ends the proof.
\end{proof}

@@

\subsection{Organization of the work}

@@

\begin{hyp}[Regularity of the mesh in 1D and constant $C_\mathcal M$] \label{geometrie1d}
 We consider that there exists 
a universal constant $0<C_{\mathcal{M}}\leq 1$ independent of the mesh size
$h=\sup_{j\in \mathbb Z} \Delta x_j $
which  controls the mesh from below
\begin{equation}\label{constante_mesh_1D}
C_{\mathcal{M}} h \leq \Delta x_j \leq h \qquad \forall j \in \mathbb{Z}.
\end{equation}
\end{hyp}

@@

The  estimates (\ref{estiuepsflux}-\ref{estiuepsflux2}) and (\ref{l2}) characterize the dissipation rate 
of both schemes.

@@

We  use the method introduced by C.

@@

that we must bound in (\ref{esti_p_h_0_p_0})  is
